---
title: "Reanalysis of ARTESIA Account for Competing Risks"
author: "Shao-Wei Lo and Sachin Shah"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true              # Enables table of contents
    toc_depth: 5           # Number of heading levels to include (e.g., h1 and h2)
    toc_float: true        # Makes the TOC float on the side as a sidebar
    number_sections: true  # Optional: Number the sections
    theme: readable        # Optional: Use a clean HTML theme
editor_options: 
  chunk_output_type: console
---
# Reconstruct KM/CIF Curves and Estimate HR & Log-Rank for Validation
The goal of this step is to demonstrate that the reconstructed event-level data set reproduces the published figure and results. 

```{r, include=FALSE}
#v3 includes looped bootstraps
rm(list=ls())

library(tidyverse)
library(survival)
library(cmprsk)
library(ggplot2)
library(cmprsk)
library(tibble)
library(knitr)
library(gt)
library(cowplot)

data <- read_csv("ipd_main.csv")
```


## Reconstruct CIF Curves from 2024 NEJM paper
This step will reproduce the inverse KM plot from Figure 1, N Engl J Med 2024;390:107-17 (https://www.nejm.org/doi/pdf/10.1056/NEJMoa2310234). This analysis treated deaths and 24 hr AF events as censoring events. 

```{r, echo = FALSE}

# Fit KM survival curves for each treatment arm
# 'status == 1' indicates the outcome event of interest (stroke/SE)
fit_cif_treat1 <- survfit(Surv(time, status == 1) ~ 1, data = data %>% filter(treat == 1), conf.type = "plain")
fit_cif_treat0 <- survfit(Surv(time, status == 1) ~ 1, data = data %>% filter(treat == 0), conf.type = "plain")

# Convert KM survival probabilities to cumulative incidence (1 - survival)
cif_data <- bind_rows(
  data.frame(
    time = fit_cif_treat1$time,
    cif = 1 - fit_cif_treat1$surv,
    group = "Apixaban (Kaplan-Meier estimator)"
  ),
  data.frame(
    time = fit_cif_treat0$time,
    cif = 1 - fit_cif_treat0$surv,
    group = "Aspirin (Kaplan-Meier estimator)"
  )
)

# Plot the reconstructed Kapalan-Meier curve
cif_ARTESIA <-
  ggplot(cif_data, aes(x = time, y = cif*100, color = group)) +
  geom_step(linewidth = 1.2) +
  labs(
    x = "Years since Randomization",
    y = "Cumulative Incidence (%)",
    color = "Treatment Group"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_blank(),  
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    panel.grid = element_blank(),  
    axis.line = element_line(linewidth = 0.8, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),  
    axis.ticks = element_line(linewidth = 0.8)  
  ) +
  scale_y_continuous(
    breaks = seq(0, 10, by = 1),  
    labels = seq(0, 10, by = 1),
    limits = c(0, 10),  
    expand = c(0, 0)  
  ) +
  scale_x_continuous(
    breaks = seq(0, 6, by = 1),  
    labels = seq(0, 6, by = 1),
    limits = c(0,NA),  
    expand = c(0, 0)  
  ) +
  scale_color_manual(values = c("Apixaban (Kaplan-Meier estimator)" = "black", "Aspirin (Kaplan-Meier estimator)" = "red"))

cif_ARTESIA
```

## Log-Rank Test for Comparison
```{r}
text_log.rank <- survdiff(Surv(time, status) ~ treat, data=data, rho=0)
text_log.rank
```
* Reconstructed: p = `r round(text_log.rank$pvalue,3)`
* Published ARTESIA: p = 0.007 


## Cox Proportional Hazards Model (cause-specific HR)
Below we estimate the cause-specific HR using the extracted data and find that it is nearly identical to the cause-specific HR reported in the trial. 
```{r}
# Fit a Cox model to estimate HR of apixaban vs. aspirin
cox_model <- coxph(Surv(time, status == 1) ~ treat, data = data)

# Extract HR and 95% CI from model summary
hr_summary <- summary(cox_model)
hr <- hr_summary$conf.int[, "exp(coef)"]  # Hazard Ratio
hr_lower <- hr_summary$conf.int[, "lower .95"]  # Lower 95% CI
hr_upper <- hr_summary$conf.int[, "upper .95"]  # Upper 95% CI

hr_results <- data.frame(
  HR = round(hr,2),
  CI_lower = round(hr_lower,2),
  CI_upper = round(hr_upper,2)
)

print(hr_results) 
```
* Reconstructed: HR = `r paste0(hr_results$HR, " (", hr_results$CI_lower, " to ", hr_results$CI_upper,")")`
* Reported by ARTESIA: HR = 0.63 (0.45 to 0.88) 


# Prepare Data for Competing Risk Analysis
The extracted event-level data identifies the time points when outcome events occur (i.e., status == 1) and when censoring events occur (i.e., status == 0). According to the primary publication, the following event types were censored: death, 24-hr AF, drop out, end of study. For this analysis we want to estimate the absolute risk reduction when death and 24 hour AF events are treated as competing events. Since we do not know the specific times when competing events (death, 24 hr AF) occurred, we probabilistically re-categorized censoring event as competing events, using the steps below. 

## Bin Follow-Up Time into Annual Intervals
We estimate the probability that an event that was recorded as a censoring event was actually a competing event on the year-level. That is, we collapse follow-up times into one-year bins (e.g., year 1, 2, ..., 6). This avoids the issue of sparse censoring events at exact time points.

```{r}
# Function to bin each discrete time point into 1 year intervals. max_time: upper bound of follow-up (default 6 years)
bin_time <- function(data, max_time = 6, interval = 1) {
  data %>%
    mutate(bin_time = ceiling(pmin(time, max_time) / interval))
}

data_bin <- bin_time(data)
```

## Summarize Censoring and Outcome Events by Year & Treatment
* status == 0 → censored: due to death, 24 hr AF, lost to follow-up, study end before a primary outcome event
* status == 1 → primary event occurred: stroke or systemic embolism
* bin_time (1–6) → Represents the year of follow-up: e.g., bin_time == 2 means Year 2
This preprocessing is essential before probabilistically reclassifying censored cases into competing risks in the next step of the analysis.

### Count censored events (status == 0) by year and treatment
```{r}
censored <- data_bin %>%
  filter(status == 0) %>%                
  group_by(bin_time, treat) %>%          
  summarise(count = n(), .groups = "drop")  

censored %>% gt() %>%  cols_label(
    bin_time = "Study year",
    treat = "Treatment arm (1 = apix, 0 = asa)",
    count = "Count of censoring events"
  )

```

### Count primary outcome events (status == 1) by year and treatment
```{r}
event <- data_bin %>%
  filter(status == 1) %>%               
  group_by(bin_time, treat) %>%         
  summarise(count = n(), .groups = "drop")

event %>% gt() %>%  cols_label(
    bin_time = "Study year",
    treat = "Treatment arm (1 = apix, 0 = asa)",
    count = "Count of primary outcome events"
  )
```


## Estimate Number of Competing Events from Trial Summary Statistics
### Create year-level risk of death and 24 hrs AF 
* Start with the number of participants at risk each year. 
* Estimate expected deaths and 24 hr AF events in a given year
* The number of competing evnets events (deaths and 24 hrs events) in a given divided by the number of all censoring events in that year gives the probability that censoring event should be recoded as a competing event.

```{r}
# Trial-reported counts from ARTESIA for each arm (N Engl J Med 2024;390:107-17)
asa_enrolled <- 1997  # aspirin arm N
asa_deaths <- 341     # deaths in aspirin arm
asa_af <- 476         # >24h AF in aspirin arm

apix_enrolled <- 2015 # apixaban arm N
apix_deaths <- 362    # Deaths in apixaban arm
apix_af <- 490        # >24h AF in apixaban arm

##Total the number at risk starting each year. These numbers are from the Figure 1
asa_at_risk <- c(1997, 1777, 1539, 1120, 780, 468)
apix_at_risk <- c(2015, 1786, 1558, 1157, 820, 474)

## censoring events extracted from inverse KM curve
censoring_asa <- censored %>% filter(treat==0) %>% select(count) %>% rename(censoring = count)
censoring_apix <- censored %>% filter(treat==1) %>% select(count) %>% rename(censoring = count)

## outcome events extracted from inverse KM curve
events_asa <- event %>% filter(treat==0) %>% select(count) %>% rename(outcome_events = count)
events_apix <- event %>% filter(treat==1) %>% select(count) %>% rename(outcome_events = count)

## create table for aspirin group
asa_df <- tibble(
  year = paste0(0:5, "-", 1:6),
  at_risk = asa_at_risk,
  )

asa_df <- bind_cols(asa_df, censoring_asa, events_asa)

##create table for apixaban group
apix_df <- tibble(
  year = paste0(0:5, "-", 1:6),
  at_risk = apix_at_risk,
  )

apix_df <- bind_cols(apix_df, censoring_apix, events_apix)

##calculate probability of 24 hr AF assuming uniform annual risk
asa_total_at_risk  <- sum(asa_at_risk)
apix_total_at_risk <- sum(apix_at_risk)

asa_df$uniform_af_rate <- asa_af/asa_total_at_risk
apix_df$uniform_af_rate <- apix_af/apix_total_at_risk

##calculate probability of death assuming a Gompertz distribution with a beta or c of 0.12
##estimate the number of competing events
##estimate the probability that an observed censoring event in the extracted data is a competing event
##first for ASA arm and then for apix arm
c = 1.12
weights = c(c^0, c^1, c^2, c^3, c^4, c^5)

###asprin arm
asa_df$weights = weights
asa_df$base_death_rate = asa_deaths / sum(asa_at_risk * weights) #base death rate account for increasing mortality rate over time
asa_df <- mutate(
  asa_df,
  weighted_death_prob = base_death_rate * weights, # weighted probability of death
  est_deaths = at_risk * weighted_death_prob, #estimated deaths 
  est_AF = at_risk * uniform_af_rate, #estimated 24 hr AF events 
  competing_events = est_deaths + est_AF, #estimated total number of competing events
  revised_censoring = censoring - competing_events, #new number of censoring events accounting for competing events
  prob_competing_events = 1 - revised_censoring / censoring #probability that an observed censoring event was a competing event 
  )

###apixaban arm
apix_df$weights = weights
apix_df$base_death_rate = apix_deaths / sum(apix_at_risk * weights)
apix_df <- mutate(
  apix_df,
  weighted_death_prob = base_death_rate * weights,
  est_deaths = at_risk * weighted_death_prob, 
  est_AF = at_risk * uniform_af_rate,
  competing_events = est_deaths + est_AF, 
  revised_censoring = censoring - competing_events, 
  prob_competing_events = 1 - revised_censoring / censoring
  )

```

```{r, echo=FALSE}
### PRINT
asa_df %>%
  select(
    year, at_risk, censoring, outcome_events, base_death_rate,
    weighted_death_prob, est_deaths, uniform_af_rate, est_AF,
    competing_events, revised_censoring, prob_competing_events
  ) %>% 
  gt() %>% 
   fmt_integer(
    columns = c(est_deaths, est_AF, competing_events, revised_censoring)
  ) %>% 
  fmt_percent(
    columns = c(base_death_rate,weighted_death_prob, uniform_af_rate),
    decimals = 2) %>%
  fmt_percent(
    columns = prob_competing_events,
    decimals = 1) %>%
  cols_label(year = md("**Year (A)**"), 
             at_risk = md("**Number at risk (B)**"), 
             censoring = md("**Total censoring events (C)**"),
             outcome_events = md("**Primary outcome count (D)**"),
             base_death_rate = md("**Death rate if the rate is uniform (E)**"),
             weighted_death_prob = md("**Death rate weighted by Gompertz distribution (F)**"),
             est_deaths = md("**Estimated number of death (G = B*F)**"), 
             uniform_af_rate = md("**24hr AF event rate, uniform (H)**"), 
             est_AF = md("**Estimated number of 24hr AF events (I = B*H)**"), 
             competing_events = md("**Total number of competing events (J = G+I)**"), 
             revised_censoring = md("**Revised number of censoring events (K = C-J)**"), 
             prob_competing_events = md("**Probability that an original coded censoring event was a competing event (L = 1-K/C)**")
             ) %>% 
  tab_header(md("**Aspirin arm**")) %>% 
  tab_options(footnotes.marks = "standard") %>% 
  tab_footnote(
    footnote = md("Values from extracted event-level data from inverse-KM curve"),
    locations = cells_column_labels(columns = c(at_risk, censoring, outcome_events))
  ) %>% 
    tab_footnote(
    footnote = md("Value based on published mortality rate by arm"),
    locations = cells_column_labels(columns = c(base_death_rate))
  ) %>% 
    tab_footnote(
    footnote = md("Value based on published 24-hr AF event rate by arm"),
    locations = cells_column_labels(columns = c(uniform_af_rate))
  ) 


apix_df %>%
  select(
    year, at_risk, censoring, outcome_events, base_death_rate,
    weighted_death_prob, est_deaths, uniform_af_rate, est_AF,
    competing_events, revised_censoring, prob_competing_events
  ) %>% 
  gt() %>% 
   fmt_integer(
    columns = c(est_deaths, est_AF, competing_events, revised_censoring)
  ) %>% 
  fmt_percent(
    columns = c(base_death_rate,weighted_death_prob, uniform_af_rate),
    decimals = 2) %>%
  fmt_percent(
    columns = prob_competing_events,
    decimals = 1) %>%
  cols_label(year = md("**Year (A)**"), 
             at_risk = md("**Number at risk (B)**"), 
             censoring = md("**Total censoring events (C)**"),
             outcome_events = md("**Primary outcome count (D)**"),
             base_death_rate = md("**Death rate if the rate is uniform (E)**"),
             weighted_death_prob = md("**Death rate weighted by Gompertz distribution (F)**"),
             est_deaths = md("**Estimated number of death (G = B*F)**"), 
             uniform_af_rate = md("**24hr AF event rate, uniform (H)**"), 
             est_AF = md("**Estimated number of 24hr AF events (I = B*H)**"), 
             competing_events = md("**Total number of competing events (J = G+I)**"), 
             revised_censoring = md("**Revised number of censoring events (K = C-J)**"), 
             prob_competing_events = md("**Probability that an original coded censoring event was a competing event (L = 1-K/C)**")
             ) %>% 
  tab_header(md("**Apixaban arm**")) %>% 
  tab_options(footnotes.marks = "standard") %>% 
  tab_footnote(
    footnote = md("Values from extracted event-level data from inverse-KM curve"),
    locations = cells_column_labels(columns = c(at_risk, censoring, outcome_events))
  ) %>% 
    tab_footnote(
    footnote = md("Value based on published mortality rate by arm"),
    locations = cells_column_labels(columns = c(base_death_rate))
  ) %>% 
    tab_footnote(
    footnote = md("Value based on published 24-hr AF event rate by arm"),
    locations = cells_column_labels(columns = c(uniform_af_rate))
  ) 


```


### Create Function 'modify_fstatus()' to Reclassifies a Portion of Censored Events as Competing Events
```{r}

competing_events_count = c(round(asa_df$competing_events),round(apix_df$competing_events))

# Function to reclassify fstatus 
modify_fstatus <- function(data_bin) {
  data_bin <- data_bin %>%
    mutate(fstatus = status)
  
  # number of cases to switch per group and bin_time
  switch_counts <- data.frame(
    treat = rep(0:1, each = 6),
    bin_time = rep(1:6, 2),
    switch_n = competing_events_count
  )
  
  # reclassify fstatus randomly
  for (i in 1:nrow(switch_counts)) {
    subset_indices <- which(data_bin$treat == switch_counts$treat[i] & 
                              data_bin$bin_time == switch_counts$bin_time[i] & 
                              data_bin$fstatus == 0)
    
    if (length(subset_indices) >= switch_counts$switch_n[i]) {
      switch_indices <- sample(subset_indices, switch_counts$switch_n[i])
      data_bin$fstatus[switch_indices] <- 2
    }
  }
  
  return(data_bin)
}

```




# Main Analysis
## Plotting Simulated CIF Curves Using Fine–Gray Subdistribution Hazard Model
### Step 1: create a CIF function
```{r}
# function to compute CIF using the 'cuminc' function
compute_cif <- function(data_bin) {
  cif <- cuminc(ftime = data_bin$time, fstatus = data_bin$fstatus, group = data_bin$treat)
  return(cif)
}
```


### Step 2: Run 1000 simulations probabalistically reclassifying competing events
```{r}
# Loop over simulations
num_simulations <- 1000
set.seed(51)

# Pre-allocate storage lists
cif_treat_0_list <- list()
cif_treat_1_list <- list()
data.m <- list()

# Pre-allocate vectors to store simulation results
risk_diffs6      <- numeric(num_simulations)
se_risk_diffs6   <- numeric(num_simulations)

cuminc_treat1_6  <- numeric(num_simulations)
cuminc_treat0_6  <- numeric(num_simulations)
se_cuminc_treat1 <- numeric(num_simulations)
se_cuminc_treat0 <- numeric(num_simulations)


# Run simulations
for (i in 1:num_simulations) {
  # create modified dataset
  modified_data_bin <- modify_fstatus(data_bin)
  data.m[[i]] <- modified_data_bin %>% select("time", "fstatus", "treat")
  
  # compute CIF
  cif <- compute_cif(modified_data_bin)
  
  # store CIF results for each treatment group
  cif_treat_0_list[[i]] <- data.frame(
    time = cif$`0 1`$time,
    cif = cif$`0 1`$est,
    simulation = i
  )
  cif_treat_1_list[[i]] <- data.frame(
    time = cif$`1 1`$time,
    cif = cif$`1 1`$est,
    simulation = i
  )
  
}
```


### Step 3: Combine simulation results
```{r}
# Combine CIF data from all simulations
cif_treat_0_all <- do.call(rbind, cif_treat_0_list)
cif_treat_1_all <- do.call(rbind, cif_treat_1_list)

# Summarize percentiles for each time point by treatment group
percentiles_cif_treat_0 <- cif_treat_0_all %>%
  group_by(time) %>%
  summarise(
    p2.5 = quantile(cif, 0.025, na.rm = TRUE),
    p50 = quantile(cif, 0.5, na.rm = TRUE),
    p97.5 = quantile(cif, 0.975, na.rm = TRUE)
  ) %>%
  mutate(group = "Aspirin (Aalen-Johansen estimator)")

percentiles_cif_treat_1 <- cif_treat_1_all %>%
  group_by(time) %>%
  summarise(
    p2.5 = quantile(cif, 0.025, na.rm = TRUE),
    p50 = quantile(cif, 0.5, na.rm = TRUE),
    p97.5 = quantile(cif, 0.975, na.rm = TRUE)
  ) %>%
  mutate(group = "Apixaban (Aalen-Johansen estimator)")

# Combine both treatment groups into one dataset
percentiles_cif_data <- bind_rows(percentiles_cif_treat_0, percentiles_cif_treat_1)
```

## Cumulative risk at 3.5 years 
* Estimate the 3.5-year cumulative risk of the primary event (stroke/systemic embolism)
* Stratify by arm and by estimator
* Calculate the differences in estimates using the KM estimator (death/24-hr AF are censored) vs. AJ estimator (death/24-hour AF are competing events)
* 95% CI are bootstrapped to account for covariance in estimates

```{r cumulative risk at 3.5 yrs, echo = FALSE}
#givens
tstar   <- 3.5      #estimation time point
cause   <- 1      #outcome code in fstatus
cencode <- 0      #censoring code in fstatus
B       <- 1000   #num of bootstraps
set.seed(51)


res <- list()

for (i in 1:num_simulations) {


#divide data by arm
dat <- data.m[[i]]
d0 <- subset(dat, treat == 0, select = c(time, fstatus))
d1 <- subset(dat, treat == 1, select = c(time, fstatus))

# Point estimates by arm 

## ---------- estimate for arm 0 ----------
# KM (competing -> censored)
s01_0 <- as.integer(d0$fstatus == cause)
sf0 <- survfit(Surv(time, s01_0) ~ 1, data = d0)
pKM0 <- 1 - summary(sf0, times = tstar, extend = TRUE)$surv

# AJ (CIF) – pick element whose name ENDS with the cause code (handles "1 1"/"1 2" naming)
ci0 <- cuminc(ftime = d0$time, fstatus = d0$fstatus, cencode = cencode)
nms0 <- names(ci0); last0 <- trimws(sapply(strsplit(nms0, " "), function(z) tail(z, 1)))
ix0 <- which(last0 == as.character(cause))[1]
  comp0 <- ci0[[ix0]]
  pAJ0 <- approx(comp0$time, comp0$est, xout = tstar,
                 method = "constant", f = 1, rule = 2, ties = "ordered")$y

## ---------- estimate for arm 1 ----------
s01_1 <- as.integer(d1$fstatus == cause)
sf1 <- survfit(Surv(time, s01_1) ~ 1, data = d1)
pKM1 <- 1 - summary(sf1, times = tstar, extend = TRUE)$surv

ci1 <- cuminc(ftime = d1$time, fstatus = d1$fstatus, cencode = cencode)
nms1 <- names(ci1); last1 <- trimws(sapply(strsplit(nms1, " "), function(z) tail(z, 1)))
ix1 <- which(last1 == as.character(cause))[1]
  comp1 <- ci1[[ix1]]
  pAJ1 <- approx(comp1$time, comp1$est, xout = tstar,
                 method = "constant", f = 1, rule = 2, ties = "ordered")$y
  


# COVARIANCE between estimators so that we can accurately calcualte the SE for the difference in estimates
## ---------- BOOTSTRAP (paired, within arm; captures KM–AJ covariance) ----------
KM0 <- AJ0 <- numeric(B)
KM1 <- AJ1 <- numeric(B)
n0 <- nrow(d0); n1 <- nrow(d1)

for (b in seq_len(B)) {
  ## Arm 0
  i0 <- sample.int(n0, n0, replace = TRUE); d0b <- d0[i0, , drop = FALSE]
  s01_0b <- as.integer(d0b$fstatus == cause)
  sf0b <- survfit(Surv(time, s01_0b) ~ 1, data = d0b)
  KM0[b] <- 1 - summary(sf0b, times = tstar, extend = TRUE)$surv

  ci0b <- try(cuminc(d0b$time, d0b$fstatus, cencode = cencode), silent = TRUE)
  if (!inherits(ci0b, "try-error")) {
    nms0b <- names(ci0b); last0b <- trimws(sapply(strsplit(nms0b, " "), function(z) tail(z, 1)))
    ix0b <- which(last0b == as.character(cause))[1]
    if (!is.na(ix0b)) {
      comp0b <- ci0b[[ix0b]]
      val <- approx(comp0b$time, comp0b$est, xout = tstar,
                    method = "constant", f = 1, rule = 2, ties = "ordered")$y
      if (is.na(val)) val <- tail(comp0b$est, 1)
      AJ0[b] <- val
    } else AJ0[b] <- 0
  } else AJ0[b] <- 0

  ## Arm 1
  i1 <- sample.int(n1, n1, replace = TRUE); d1b <- d1[i1, , drop = FALSE]
  s01_1b <- as.integer(d1b$fstatus == cause)
  sf1b <- survfit(Surv(time, s01_1b) ~ 1, data = d1b)
  KM1[b] <- 1 - summary(sf1b, times = tstar, extend = TRUE)$surv

  ci1b <- try(cuminc(d1b$time, d1b$fstatus, cencode = cencode), silent = TRUE)
  if (!inherits(ci1b, "try-error")) {
    nms1b <- names(ci1b); last1b <- trimws(sapply(strsplit(nms1b, " "), function(z) tail(z, 1)))
    ix1b <- which(last1b == as.character(cause))[1]
    if (!is.na(ix1b)) {
      comp1b <- ci1b[[ix1b]]
      val <- approx(comp1b$time, comp1b$est, xout = tstar,
                    method = "constant", f = 1, rule = 2, ties = "ordered")$y
      if (is.na(val)) val <- tail(comp1b$est, 1)
      AJ1[b] <- val
    } else AJ1[b] <- 0
  } else AJ1[b] <- 0
}

# OUTPUT
D0 <- AJ0 - KM0
D1 <- AJ1 - KM1

est_arm0 <- c(KM = pKM0, AJ = pAJ0, Diff = pAJ0 - pKM0)
est_arm1 <- c(KM = pKM1, AJ = pAJ1, Diff = pAJ1 - pKM1)

ci_arm0 <- quantile(D0, c(0.025, 0.975), na.rm = TRUE)
ci_arm1 <- quantile(D1, c(0.025, 0.975), na.rm = TRUE)

cov_arm0 <- cov(KM0, AJ0); cor_arm0 <- cov_arm0 / (sd(KM0) * sd(AJ0))
cov_arm1 <- cov(KM1, AJ1); cor_arm1 <- cov_arm1 / (sd(KM1) * sd(AJ1))

res[[i]] <- data.frame(
  arm = c("Aspirin", "Apixaban"),
  KM_risk = c(est_arm0["KM"],  est_arm1["KM"]),
  AJ_risk = c(est_arm0["AJ"],  est_arm1["AJ"]),
  Diff_AJ_minus_KM = c(est_arm0["Diff"], est_arm1["Diff"]),
  CI_low = c(ci_arm0[1], ci_arm1[1]),
  CI_high = c(ci_arm0[2], ci_arm1[2])
  #KM_AJ_cov = c(cov_arm0, cov_arm1),
  #KM_AJ_cor = c(cor_arm0, cor_arm1)
)
print(i)
}

KM.asa <- numeric(num_simulations)
KM.apix<- numeric(num_simulations)
AJ.asa<- numeric(num_simulations)
AJ.apix<- numeric(num_simulations)
diff.asa<- numeric(num_simulations)
diff.apix<- numeric(num_simulations)
ci_low_apix <- numeric(num_simulations)
ci_low_asa <- numeric(num_simulations)
ci_high_apix <- numeric(num_simulations)
ci_high_asa <- numeric(num_simulations)

for (i in 1:num_simulations) {
  KM.asa <- res[[i]][1,2]
  KM.apix<- res[[i]][2,2]
  AJ.asa<- res[[i]][1,3]
  AJ.apix<- res[[i]][2,3]
  diff.asa<- res[[i]][1,4]
  diff.apix<- res[[i]][2,4]
  ci_low_asa <- res[[i]][1,5]
  ci_low_apix <- res[[i]][2,5]
  ci_high_asa <- res[[i]][1,6]
  ci_high_apix <- res[[i]][2,6]
}

quantile(ci_low_apix, 0.5)
quantile(ci_high_apix, 0.5)
quantile(ci_low_asa, 0.5)
quantile(ci_high_asa, 0.5)

pct <- function(x) sprintf("%.2f%%", 100*x)

res_print <- 
  data.frame(
    arm=c("Aspirin", "Apixaban"), 
    KM_risk = c(pct(quantile(KM.asa, 0.5)), pct(quantile(KM.apix, 0.5))),
    AJ_risk = c(pct(quantile(AJ.asa, 0.5)), pct(quantile(AJ.apix, 0.5))),
    Diff_AJ_minus_KM = c(pct(quantile(diff.asa, 0.5)), pct(quantile(diff.apix, 0.5))),
    CI = c(
      paste0(pct(quantile(ci_low_asa,0.5)), " to ", pct(quantile(ci_high_asa,0.5))),
      paste0(pct(quantile(ci_low_apix,0.5)), " to ", pct(quantile(ci_high_apix,0.5)))
    )
  )

res_print %>% gt() %>%  cols_label(
  arm = md("**Group**"),  
  KM_risk = md("**Risk with KM estimator**"),
    AJ_risk = md("**Risk with AJ estimator**"),
    Diff_AJ_minus_KM = md("**Difference**"),
    CI = md("**95% confidence interval**")
  ) 


```
	
## Risk Difference at 3.5 Years using KM and AJ estimators
* We estiamte the ARR using the KM estimator and the AJ estimator
* We take the difference and use bootstrapping to define the 95% CI

```{r need to change, echo = FALSE}

## --- SETTINGS ---
tstar   <- 3.5      # horizon (same units as dat$time)
cause   <- 1      # event-of-interest code in fstatus
cencode <- 0      # censoring code
B       <- 1000
set.seed(51)

pval_Delta <- numeric(num_simulations)
RD_KM_hat<- numeric(num_simulations)
RD_AJ_hat<- numeric(num_simulations)
Delta_hat<- numeric(num_simulations)
ci_RD_KM_low<- numeric(num_simulations)
ci_RD_KM_high<- numeric(num_simulations)
ci_RD_AJ_low<- numeric(num_simulations)
ci_RD_AJ_high<- numeric(num_simulations)
ci_Delta_low<- numeric(num_simulations)
ci_Delta_high<- numeric(num_simulations)


for (i in 1:num_simulations) {
  
dat <- data.m[[i]]
d0 <- subset(dat, treat == 0, select = c(time, fstatus))
d1 <- subset(dat, treat == 1, select = c(time, fstatus))

## --- POINT ESTIMATES ---
# Arm 0 KM
s01_0 <- as.integer(d0$fstatus == cause)
sf0   <- survfit(Surv(time, s01_0) ~ 1, data = d0)
KM0   <- 1 - summary(sf0, times = tstar, extend = TRUE)$surv

# Arm 0 AJ
ci0   <- cuminc(ftime = d0$time, fstatus = d0$fstatus, cencode = cencode)
nms0  <- names(ci0); last0 <- trimws(sapply(strsplit(nms0, " "), function(z) tail(z, 1)))
ix0   <- which(last0 == as.character(cause))[1]
#if (!is.na(ix0)) {
  comp0 <- ci0[[ix0]]
  AJ0   <- approx(comp0$time, comp0$est, xout = tstar,
                  method = "constant", f = 1, rule = 2, ties = "ordered")$y
#  if (is.na(AJ0)) AJ0 <- tail(comp0$est, 1)
#} else AJ0 <- 0

# Arm 1 KM
s01_1 <- as.integer(d1$fstatus == cause)
sf1   <- survfit(Surv(time, s01_1) ~ 1, data = d1)
KM1   <- 1 - summary(sf1, times = tstar, extend = TRUE)$surv

# Arm 1 AJ
ci1   <- cuminc(ftime = d1$time, fstatus = d1$fstatus, cencode = cencode)
nms1  <- names(ci1); last1 <- trimws(sapply(strsplit(nms1, " "), function(z) tail(z, 1)))
ix1   <- which(last1 == as.character(cause))[1]
#if (!is.na(ix1)) {
  comp1 <- ci1[[ix1]]
  AJ1   <- approx(comp1$time, comp1$est, xout = tstar,
                  method = "constant", f = 1, rule = 2, ties = "ordered")$y
#  if (is.na(AJ1)) AJ1 <- tail(comp1$est, 1)
#} else AJ1 <- 0

RD_KM_hat[[i]] <- KM1 - KM0
RD_AJ_hat[[i]] <- AJ1 - AJ0
Delta_hat[[i]] <- (AJ1 - AJ0) - (KM1 - KM0)


## --- BOOTSTRAP (paired within arms) ---
n0 <- nrow(d0); n1 <- nrow(d1)
RD_KM <- RD_AJ <- Delta <- numeric(B)

for (b in seq_len(B)) {
  ## resample rows within each arm
  i0 <- sample.int(n0, n0, replace = TRUE); d0b <- d0[i0, , drop = FALSE]
  i1 <- sample.int(n1, n1, replace = TRUE); d1b <- d1[i1, , drop = FALSE]

  ## KM arm 0
  s01_0b <- as.integer(d0b$fstatus == cause)
  sf0b   <- survfit(Surv(time, s01_0b) ~ 1, data = d0b)
  KM0b   <- 1 - summary(sf0b, times = tstar, extend = TRUE)$surv

  ## KM arm 1
  s01_1b <- as.integer(d1b$fstatus == cause)
  sf1b   <- survfit(Surv(time, s01_1b) ~ 1, data = d1b)
  KM1b   <- 1 - summary(sf1b, times = tstar, extend = TRUE)$surv

  ## AJ arm 0
  ci0b <- try(cuminc(d0b$time, d0b$fstatus, cencode = cencode), silent = TRUE)
  if (!inherits(ci0b, "try-error")) {
    n0n  <- names(ci0b); last0b <- trimws(sapply(strsplit(n0n, " "), function(z) tail(z, 1)))
    j0   <- which(last0b == as.character(cause))[1]
    if (!is.na(j0)) {
      c0b <- ci0b[[j0]]
      AJ0b <- approx(c0b$time, c0b$est, xout = tstar,
                     method = "constant", f = 1, rule = 2, ties = "ordered")$y
      if (is.na(AJ0b)) AJ0b <- tail(c0b$est, 1)
    } else AJ0b <- 0
  } else AJ0b <- 0

  ## AJ arm 1
  ci1b <- try(cuminc(d1b$time, d1b$fstatus, cencode = cencode), silent = TRUE)
  if (!inherits(ci1b, "try-error")) {
    n1n  <- names(ci1b); last1b <- trimws(sapply(strsplit(n1n, " "), function(z) tail(z, 1)))
    j1   <- which(last1b == as.character(cause))[1]
    if (!is.na(j1)) {
      c1b <- ci1b[[j1]]
      AJ1b <- approx(c1b$time, c1b$est, xout = tstar,
                     method = "constant", f = 1, rule = 2, ties = "ordered")$y
      if (is.na(AJ1b)) AJ1b <- tail(c1b$est, 1)
    } else AJ1b <- 0
  } else AJ1b <- 0

  ## store contrasts
  RD_KM[b] <- KM1b - KM0b
  RD_AJ[b] <- AJ1b - AJ0b
  Delta[b] <- RD_AJ[b] - RD_KM[b]
}

## --- CIs (percentile; paired by construction) ---
ci_RD_KM  <- quantile(RD_KM, c(0.025, 0.975), na.rm = TRUE)
ci_RD_AJ  <- quantile(RD_AJ, c(0.025, 0.975), na.rm = TRUE)
ci_Delta  <- quantile(Delta, c(0.025, 0.975), na.rm = TRUE)

ci_RD_KM_low[[i]] <- ci_RD_KM[1]
ci_RD_KM_high[[i]] <- ci_RD_KM[2]
ci_RD_AJ_low[[i]] <- ci_RD_AJ[1]
ci_RD_AJ_high[[i]] <- ci_RD_AJ[2]
ci_Delta_low[[i]]<- ci_Delta[1]
ci_Delta_high[[i]] <- ci_Delta[2]

## --- Bootstrap p-value for H0: Delta = 0 (symmetric two-sided) ---
pval_Delta[[i]] <- 2 * min(
  mean(Delta >= 0, na.rm = TRUE),
  mean(Delta <= 0, na.rm = TRUE)
)
print(i)
}

## --- PRINT (percent format) ---
pct <- function(x) sprintf("%.2f%%", 100*x)

data.diff <- data.frame(
  Estimators = c("Kaplan-Meier", "Aalen-Johansen", "Difference"),
  Event = c("Censoring events", "Competing events", "--"),
  Estimate = c(
    pct(quantile(RD_KM_hat, 0.5)),
    pct(quantile(RD_AJ_hat, 0.5)),
    pct(quantile(Delta_hat, 0.5))
  ),
  CI = c(paste0(
          pct(quantile(ci_RD_KM_low, 0.5)), 
          " to ", 
          pct(quantile(ci_RD_KM_high, 0.5))
                ), 
        paste0(
          pct(quantile(ci_RD_AJ_low, 0.5)), 
          " to ", 
          pct(quantile(ci_RD_AJ_high, 0.5))
        ),
        paste0(
        pct(quantile(ci_Delta_low, 0.5)), 
          " to ", 
        pct(quantile(ci_Delta_high, 0.5))
              )
        )
    ) 

data.diff %>% gt() %>% 
  cols_label(
    Estimators = md("**Estimator**"),
    Event = md("**How are deaths and 24hr-AF events handled?**"),
    Estimate = md("**Estimate**"),
    CI = md("**95% Confidence interval**")
    )




```



## Plot
```{r, echo = FALSE, message = FALSE, warning=FALSE, fig.width=10, fig.height=9, fig.align='center',}
#theme

cust_min_theme <-  theme_minimal() +
  theme(
    plot.title = element_blank(),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_blank(),  
    legend.text = element_text(size = 12),
    legend.position = "top",  
    legend.box = "horizontal",  
    legend.spacing.x = unit(0.5, "cm"),  
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.8, color = "black"),
    axis.ticks.length = unit(0.3, "cm"),
    axis.ticks = element_line(linewidth = 0.8)
  ) 

plot.full <-  ggplot() +
  # Median CIF curves from competing risks simulations
  geom_line(data = percentiles_cif_data, aes(x = time, y = p50*100, color = group), linewidth = 1) +
  geom_step(data = cif_data, aes(x = time, y = cif*100, color = group), size = 1) +
   scale_y_continuous(
    breaks = seq(0, 100, by = 10),
    labels = seq(0, 100, by = 10),
    limits = c(0, 100),
    expand = c(0, 0)
  ) +
  scale_x_continuous(
    breaks = seq(0, 6, by = 1),
    labels = seq(0, 6, by = 1),
    limits = c(0, NA),
    expand = c(0, 0)
  ) +
  scale_color_manual(values = c(
    "Apixaban (Kaplan-Meier estimator)" = "grey",
    "Aspirin (Kaplan-Meier estimator)" = "pink",
    "Apixaban (Aalen-Johansen estimator)" = "black",
    "Aspirin (Aalen-Johansen estimator)" = "red"
  )) +
   labs(
    title = "Cumulative Incidence Function (CIF) by Treatment Group",
    x = "Years since Randomization",
    y = "Cumulative Incidence (%)"
    #color = "Treatment Group",
    #fill = "Treatment Group"
  ) +
   cust_min_theme +
  guides(color = guide_legend(nrow = 2, byrow = TRUE)) 


## format ARR lablels

paste0(
  data.diff[1,3], 
  " (95% CI, ",
  data.diff[1,4], 
  ")"
)

text_arr_km <- paste0(
  data.diff[1,3], 
  " (95% CI, ",
  data.diff[1,4], 
  ")"
)

text_arr_aj <-paste0(
  data.diff[2,3], 
  " (95% CI, ",
  data.diff[2,4], 
  ")"
)

text_diff <- paste0(
  data.diff[3,3], 
  " (95% CI, ",
  data.diff[3,4], 
  ")"
)


##zoomed inset

plot.zoomed <- 
  plot.full +
   scale_y_continuous(
    breaks = seq(0, 10, by = 1),
    labels = seq(0, 10, by = 1),
    limits = c(0, 10),
    expand = c(0, 0)
  ) +
  cust_min_theme +
  theme(
  legend.position="none",
  axis.title.x=element_blank(), 
  axis.title.y=element_blank()
  )  +
  annotate(geom="text", x = 0.25, y=8.8, hjust =0, label=paste0("ARR by Kaplan-Meier at 3.5 yrs.= ", text_arr_km)) + 
  annotate(geom="text", x = 0.25, y=9.2, hjust=0, label=paste0("ARR by Aalen-Johansen at 3.5 yrs. = ", text_arr_aj)) +
  annotate(geom="text", x = 0.25, y=8.4, hjust=0, label=paste0("Difference in estimates at 3.5 yrs. = ", text_diff))
  
##add risk table

risk_table <- data.frame(
  treat = rep(c("Aspirin", "Apixaban"), each = 7), 
  time = seq(0, 6, by = 1), 
  n_risk = c(asa_at_risk, 200, apix_at_risk, 214)
)

ticks <- seq(0, 6, by = 1)

p_tab <- ggplot(risk_table, aes(x = time, y = treat, label = n_risk)) +
  geom_text() +
  scale_x_continuous(breaks = ticks) +
  labs(x = NULL, y = NULL, title = "Number at risk") +
  theme_minimal(base_size = 11) +
  theme(axis.text.y = element_text(size=12, color = "black"),
        panel.grid = element_blank(),
        axis.text.x = element_blank(),
        plot.title = element_text(hjust = 0),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank()
    )
    

#align table and plot
top_wider <- plot_grid(NULL, plot.full, NULL, 
                     ncol = 3, 
                     rel_widths = c(0.052, 1, 0.04))  # middle is real plot

plot.combined <- plot_grid(top_wider, p_tab, 
                   ncol = 1 ,
          rel_heights = c(1,0.15))

plot.inset <- 
  ggdraw() +
  draw_plot(plot.combined) +
  draw_plot(plot.zoomed, x = 0.2, y = 0.35, width = 0.76, height = 0.56)


plot.inset
ggsave(paste0("CIF plot ", format(Sys.time(), "%Y-%m-%d_%H-%M"), ".pdf"), width = 10, height = 9)   
ggsave(paste0("CIF plot ", format(Sys.time(), "%Y-%m-%d_%H-%M"), ".png"), width = 10, height = 9)   

#height 900; width 1000


```
 
